---
title: "Project-draft"
output: html_document
date: '2022-06-13'
---

```{r}
library(tidyverse)
library(tidymodels)
library(tidyverse)
library(tidymodels)
library(dplyr)
library(ggplot2)
library(cowplot)
library(knitr)
library(recipes)
library(caret)
library(InformationValue)
library(ISLR)
library(MASS)
library(nnet)

```

```{r}
#| label: load data
redwine <- read.csv("data/winequality-red.csv", sep = ";")
whitewine <- read.csv("data/winequality-white.csv", sep = ";")
redwine<-redwine%>%mutate(color="red")
whitewine<-whitewine%>%mutate(color="white")
wine<-redwine%>%full_join(whitewine)
wine<- slice(wine, sample(1:n()))
```
```{r}
ggplot(data = wine, aes(x = fixed.acidity, y = quality, color = color)) +
  geom_jitter()
```

```{r}
wine<-wine%>%
  mutate(good_wine=if_else(quality >= 7,"1","0"))
wine<-wine%>%
  mutate(good_wine_names=if_else(good_wine=="1","Good wine","Bad or subpar wine"))
ggplot(wine,aes(x=good_wine))+
  geom_bar()+
  labs(title="Distribution of Good wine in data",
       y="Count",
       x="Good Wine"
  )
```
```{r}
no1 <- colnames(wine)[1:11]
colnames(wine)[1:11] = paste("c_", no1, sep = "")
```

```{r datasplit}
set.seed(222)
wine_split <- initial_split(wine, prop = 3/4)
wine_train <- training(wine_split)
wine_test <- training(wine_split)
```

```{r specmodel}
wine_spec <- logistic_reg() %>%
  set_engine("glm")
```

## Logistic Model

### Initial
```{r}
wine_rec1 <- recipe(
  good_wine ~., data = wine_train) %>%
  step_rm(good_wine_names) %>%
  step_interact(terms = ~starts_with("c_"):color) %>%
  step_rm(quality) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_zv(all_predictors())
```

```{r}
wine_wflow1 <- workflow() %>%
  add_model(wine_spec) %>%
  add_recipe(wine_rec1)
```

```{r}
wine_fit1 <- wine_wflow1 %>%
  fit(data = wine_train)
kable(tidy(wine_fit1), digits = 3)
```
We'll remove citric.acid.

### Full model
```{r}
wine_rec2 <- recipe(
  good_wine ~., data = wine_train) %>%
  step_rm(good_wine_names) %>%
  step_interact(terms = ~starts_with("c_"):color) %>%
  step_rm(quality) %>%
  step_rm(c_free.sulfur.dioxide_x_colorwhite) %>%
  step_rm(c_chlorides_x_colorwhite) %>%
  step_rm(c_citric.acid_x_colorwhite) %>%
  step_rm(c_residual.sugar_x_colorwhite) %>%
  step_rm(c_volatile.acidity_x_colorwhite) %>%
  step_rm(c_fixed.acidity_x_colorwhite) %>%
  step_rm(c_free.sulfur.dioxide) %>%
  step_rm(c_citric.acid) %>%
  step_rm(c_pH) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_zv(all_predictors())
```

```{r}
wine_wflow2 <- workflow() %>%
  add_model(wine_spec) %>%
  add_recipe(wine_rec2)
```

```{r}
wine_fit2 <- wine_wflow2 %>%
  fit(data = wine_train)
kable(tidy(wine_fit2), digits = 3)
```

$log~odds(good~wine) = 290.153 + 0.371*fixed~acidity - 3.217*volatile + 0.256* residual~sugar - 11.862*chlorides - 0.011*total.sulfur.dioxide - 304.879*density + 3.63*sulphates + 0.766*alcohol + 0.015*total.sulfur.dioxide x colorwhite - 265.692*density x colorwhite - 1.534*sulphates x colorwhite - 0.511*alcohol x colorwhite + 259.744*colorwhite$

```{r}
wine_fit_test <- wine_wflow2 %>%
  fit(data = wine_test)

tidy(wine_fit_test) %>% 
  kable(digits = 3)
```
```{r}
wine_test_pred2 <- predict(wine_fit2, wine_test, type = "prob") %>% 
  bind_cols(wine_test) 
wine_test_pred2
```

### Reduced Model (no interactive terms)
```{r}
wine_rec3 <- recipe(
  good_wine ~., data = wine_train) %>%
  step_rm(good_wine_names) %>%
  step_rm(quality) %>%
  step_rm(c_citric.acid) %>%
  step_rm(c_total.sulfur.dioxide) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_zv(all_predictors())
```

```{r}
wine_wflow3 <- workflow() %>%
  add_model(wine_spec) %>%
  add_recipe(wine_rec3)
```

```{r}
wine_fit3 <- wine_wflow3 %>%
  fit(data = wine_train)
kable(tidy(wine_fit3), digits = 3)
```

```{r}
wine_test_pred3 <- predict(wine_fit3, wine_test, type = "prob") %>% 
  bind_cols(wine_test) 
wine_test_pred3
```



### Model selection
```{r}
wine_fit2_eg <- wine_fit2 %>% extract_fit_engine()
wine_fit3_eg <- wine_fit3 %>% extract_fit_engine()
```

```{r}
wine_test_pred2 %>% roc_curve(truth = as.factor(good_wine), .pred_0) %>% mutate(model = "Model 1") %>%
  bind_rows(wine_test_pred3 %>% roc_curve(truth = as.factor(good_wine), .pred_0) %>% mutate(model = "Model 2")) %>%
  ggplot(aes(x = 1 - specificity, y = sensitivity, color = model)) +
  geom_line()
```

```{r}
wine_test_pred2 %>%
  roc_auc(truth = as.factor(good_wine), .pred_0)

wine_test_pred3 %>%
  roc_auc(truth = as.factor(good_wine), .pred_0)
```

Based on the roc_auc, the full model performs slightly better than the reduced model.

AIC
```{r}
glance(wine_fit2_eg)$AIC
glance(wine_fit3_eg)$AIC
```
BIC
```{r}
glance(wine_fit2_eg)$BIC
glance(wine_fit3_eg)$BIC
```

The full model has lower AIC and BIC

```{r}
anova(wine_fit2_eg, wine_fit3_eg)
```
Drop-in-Deviance Test
```{r}
pchisq(38.508, 4, lower.tail = FALSE)
```
The p-value is very small, smaller than the critical value of 0.05 under 95% CI. So we can reject the null hypothesis and conclude that there is enough evidence showing that there is at least 1 beta_j does not equal to 0.

Conclusion: we'll use the full model.
```{r}
wine2<-wine%>%mutate(quality=factor(quality,levels=0:10))
```
```{r}
set.seed(22)
wine_split2 <- initial_split(wine2, prop = 3/4)
wine_train <- training(wine_split2)
wine_test <- testing(wine_split2)
```

```{r}
full_fit1<- multinom_reg() %>%
  set_engine("nnet") %>%
  fit(quality~.,
  data = wine_train)
```
```{r}
full_fit2<- repair_call(full_fit2, data = wine_train)
full_fit2_fixed<-full_fit2 %>% extract_fit_engine() 
```
```{r}
newmodel<-stepAIC(full_fit2_fixed,direction="both")
```

```{r}
full_fit2%>%tidy()
newmodel%>%tidy()
newmodel$AIC
glance(full_fit1)$AIC
```
```{r}
training_pred <- predict(full_fit2,wine_test)

accuracy <- mean(training_pred$.pred_class == wine_test$quality)
```
```{r}
training_pred$.pred_class<-newmodel%>%predict(wine_test)
training_pred2<-training_pred%>%mutate(training_pred2=factor(.pred_class,levels=0:10))

accuracy2 <- mean(training_pred2$training_pred2 == wine_test$quality)

```