---
title: "project-revised"
subtitle: "STA 210 - Summer 2022"
author: "Team 2 - Alicia Gong, Ashley Chen, Abdel Shehata, Claire Tam"
date: "6-16-2022"
format: pdf
editor: visual
editor_options: 
  markdown: 
    wrap: 72
---

### Setup

#### Load packages and data:

```{r}
#| label: load-pkg-data
#| message: false
#| warning: false
library(tidyverse)
library(tidymodels)
library(dplyr)
library(ggplot2)
library(cowplot)
library(knitr)
library(recipes)
library(caret)
library(InformationValue)
library(ISLR)
library(MASS)
library(nnet)
```

```{r}
#| label: load data

redwine <- read.csv("data/winequality-red.csv", sep = ";")
whitewine <- read.csv("data/winequality-white.csv", sep = ";")
redwine <- redwine %>% mutate(color="red")
whitewine <- whitewine %>% mutate(color="white")
wine <- redwine %>% full_join(whitewine)
wine <- slice(wine, sample(1:n()))
```

### Introduction and Data

#### Introduction

About 234 million hectoliters of wine were consumed in 2020, worldwide,
with the US making up approximetly 14% of that consumption (Karlsson
2020). Since Wine composition and wine quality varies widely, it raises
the question: what makes a good wine?

\par

To answer that question, we will analyze the wine quality dataset from
Vinho Verde vinyard in Portogal, and more importantly try to narrow down
our question to make it possible for it to be supported by evidence.
Below is the introduction to our research:

\par

Project Goal: To identify variables that are important in explaining
variation in the response. "Vinho Verde" is the kind of wine exclusively
produced in the demarcated region of Vinho Verde in northwestern
Portugal. The Vinho Verde wine has its own unique system of production
and is only produced from the indigenous grape varieties of the region.
Vinho Verde region is one of the largest and oldest wine regions in the
world, and is home to thousands of producers, generating a wealth of
economic activity and jobs, and strongly contributes to the development
of Minho province and the country. The Vinho Verde wine also enjoys high
reputation worldwide. It is recurrently awarded in national and
international competitions. The goal of this dataset is to model wine
quality based on physicochemical tests. We believe that this dataset can
also be used to analyze the relationship between different chemical
compositions and the ratings of wine quality. We believe that this
dataset can also be used to analyze what chemical factors are
attributable to the final rating of Vinho Verde wine. Our research may
shed light on future research and development directions for improving
the quality of Vinho Verde wine, which may also contribute to the
competitiveness of Portuguese wine industry.

\par

Our goal is to produce a classification model that best explains how
different chemical compositions of the Portuguese "Vinho Verde" wine
affects the variation of the wine quality.

#### Data Introduction

The Wine Quality dataset was collected from Vinho Verde wine Samples,
from the North of Portugal. The data was originally donated in 2009 by
Professor Cortez. The specific mechanism of the collection of the data
was lab work done on different wines to measure their chemical
attributites (like acidity etc.). The quality of the wine however was
obtained through the average rating of three wine experts. The dataset
is divided into two: Red wine and White wine. Red Wine has 1599
observations, and white wine has 4898 observations (each observation
being a specific wine). Information about the wine include but are not
limited to:PH,Density,Acidity, and alcohol content.

\par

Each observation is a specific wine from the Vinho Verde region. Thus,
there might be a little uncertainity in collecting the exact numbers for
each numbers. However, since the Vinho Verde region is a vast region
spreading 15500 hectareas of vineyards in far-north Portugal, this
uncertainity shouldn't be significant in our analysis or project. Thus,
we will assume that the datas are independent and random

```{r glimpse}
glimpse(wine)
```

There are 6497 observations and 13 variables (14 if you include the new
response variable added later).

```{r}
any(is.na(wine))
```

There are no NAs in our data, so we shouldn't be concerned about missing
data.

#### Data Wrangling

```{r}
wine <- wine %>%
  mutate(good_wine = if_else(quality >= 7,"1","0"))
wine <- wine %>%
  mutate(good_wine = as.factor(good_wine))
wine <- wine %>%
  mutate(good_wine_names = if_else(good_wine=="1","Good wine","Bad or subpar wine"))

no1 <- colnames(wine)[1:11]
colnames(wine)[1:11] = paste("c_", no1, sep = "")
```

#### Exploratory Data Analysis

```{r quality-dist}
ggplot(wine, aes(x = good_wine)) +
  geom_bar() +
  labs(title = "Distribution of good wine in data",
       y = "Count",
       x = "Good Wine"
  )

ggplot(wine, aes(x = quality)) +
  geom_bar() +
  labs(title = "Distribution of wine quality in data",
       y = "Count",
       x = "Quality"
  )
```

```{r distributions}
p1  <- ggplot(data = wine, aes(x = quality) ) + 
  geom_bar(fill = "pink") + 
  labs(x = "quality")

p2  <- ggplot(data = wine, aes(x = c_fixed.acidity) ) + 
  geom_histogram(fill = "pink") + 
  labs(x = "fixed acidity")

p3  <- ggplot(data = wine, aes(x = c_volatile.acidity) ) + 
  theme(axis.text=element_text(size=9)) + 
  geom_histogram(fill = "pink") + 
  labs(x = "volatile acidity")

p4  <- ggplot(data = wine, aes(x = c_citric.acid) ) + 
  theme(axis.text = element_text(size=9)) + 
  geom_histogram(fill = "pink") + 
  labs(x = "citric acid")

p5  <- ggplot(data = wine, aes(x = c_residual.sugar) ) +
  geom_histogram(fill = "pink") + 
  labs(x = "residual sugar")

p6  <- ggplot(data = wine, aes(x = c_chlorides) ) + 
  theme(axis.text = element_text(size = 11)) + 
  geom_histogram(fill = "pink") + 
  labs(x = "chlorides")

p7  <- ggplot(data = wine, aes(x = c_free.sulfur.dioxide) ) + 
  theme(axis.text = element_text(size=9)) + 
  geom_histogram(fill = "pink") + 
  labs(x = "free sulfur dioxide")

p8  <- ggplot(data = wine, aes(x = c_total.sulfur.dioxide) ) + 
  theme(axis.text = element_text(size=9)) + 
  geom_histogram(fill = "pink") + 
  labs(x = "total sulfur dioxide")

p9  <- ggplot(data = wine, aes(x = c_density) ) + 
  theme(axis.text = element_text(size = 7.5)) + 
  geom_histogram(fill= "pink") + 
  labs(x = "density")

p10 <- ggplot(data = wine, aes(x = c_pH) ) + 
  geom_histogram(fill = "pink") + 
  labs(x = "pH")

p11 <- ggplot(data = wine, aes(x = c_sulphates) ) + 
  theme(axis.text = element_text(size=9)) + 
  geom_histogram(fill= "pink") + 
  labs(x = "sulphates")

p12 <- ggplot(data = wine, aes(x = c_alcohol) ) + 
  geom_histogram(fill= "pink") + 
  labs(x = "alcohol")

plot_grid(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, p11, p12, ncol = 3, nrow = 4)
```

```{r garbage-reference}

#cor(wine$c_alcohol, wine$quality)
#cor(wine$c_density, wine$quality)
#cor(wine$c_volatile.acidity, wine$quality)
#cor(wine$c_chlorides, wine$quality)
#cor(wine$c_residual.sugar, wine$quality)
#cor(wine$c_fixed.acidity, wine$quality)
#cor(wine$c_free.sulfur.dioxide, wine$quality)
#cor(wine$c_total.sulfur.dioxide, wine$quality)
#cor(wine$c_pH, wine$quality)
#cor(wine$c_sulphates, wine$quality)
#cor(wine$c_citric.acid, wine$quality)
```

```{r}
a1 <- ggplot(wine, aes(y = c_alcohol, x = as.factor(good_wine))) + 
  geom_boxplot() + 
  labs(x = "good wine", y = "alcohol")

a2 <- ggplot(wine, aes(y = c_density, x = as.factor(good_wine))) + 
  geom_boxplot() + 
  labs(x = "good wine", y = "density")

a3 <- ggplot(wine, aes(y = c_volatile.acidity, x = as.factor(good_wine))) + 
  geom_boxplot() + 
  labs(x = "good wine", y = "volatile acidity")

a4 <- ggplot(wine, aes(y = c_chlorides, x = as.factor(good_wine))) + 
  geom_boxplot() + 
  labs(x = "good wine", y = "chlorides")

plot_grid(a1, a2, a3, a4, ncol = 2, nrow = 2)

# exploring interaction effects
a5 <- ggplot(wine, aes(y = c_sulphates, x = as.factor(good_wine), color = color)) + 
  geom_boxplot() + 
  labs(x = "good wine", y = "chlorides")

a6 <- ggplot(wine, aes(y = c_pH, x = as.factor(good_wine), color = color)) + 
  geom_boxplot() + 
  labs(x = "good wine", y = "chlorides")

a7 <- ggplot(wine, aes(y = c_total.sulfur.dioxide, x = as.factor(good_wine), color = color)) + 
  geom_boxplot() + 
  labs(x = "good wine", y = "chlorides")

a8 <- ggplot(wine, aes(y = c_alcohol, x = as.factor(good_wine), color = color)) + 
  geom_boxplot() + 
  labs(x = "good wine", y = "alcohol")

plot_grid(a5, a6, a7, a8, ncol = 2, nrow = 2)
```

```{r}
a1 <- ggplot(wine, aes(y = c_alcohol, x = as.factor(quality))) + 
  geom_boxplot() + 
  labs(x = "good wine", y = "alcohol")

a2 <- ggplot(wine, aes(y = c_density, x = as.factor(quality))) + 
  geom_boxplot() + 
  labs(x = "good wine", y = "density")

a3 <- ggplot(wine, aes(y = c_volatile.acidity, x = as.factor(quality))) + 
  geom_boxplot() + 
  labs(x = "good wine", y = "volatile acidity")

a4 <- ggplot(wine, aes(y = c_chlorides, x = as.factor(quality))) + 
  geom_boxplot() + 
  labs(x = "good wine", y = "chlorides")

plot_grid(a1, a2, a3, a4, ncol = 2, nrow = 2)

# exploring interaction effects
a5 <- ggplot(wine, aes(y = c_sulphates, x = as.factor(quality), color = color)) + 
  geom_boxplot() + 
  labs(x = "good wine", y = "chlorides")

a6 <- ggplot(wine, aes(y = c_pH, x = as.factor(quality), color = color)) + 
  geom_boxplot() + 
  labs(x = "good wine", y = "chlorides")

a7 <- ggplot(wine, aes(y = c_total.sulfur.dioxide, x = as.factor(quality), color = color)) + 
  geom_boxplot() + 
  labs(x = "good wine", y = "chlorides")

a8 <- ggplot(wine, aes(y = c_alcohol, x = as.factor(quality), color = color)) + 
  geom_boxplot() + 
  labs(x = "good wine", y = "alcohol")

plot_grid(a5, a6, a7, a8, ncol = 2, nrow = 2)
```

### Methodology

#### Initial Logistic Model

Data split: we split the data into 25% testing set and 75% training set.

```{r datasplit}
set.seed(222)

wine_split <- initial_split(wine, prop = 3/4)
wine_train <- training(wine_split)
wine_test <- training(wine_split)
```

```{r specmodel}
wine_spec <- logistic_reg() %>% set_engine("glm")
```

We plan to compare a full model and a reduced model. For the full model,
we decide to include interactive terms. For the reduced model, we adopt the stepwise AIC test to get a best AIC model. 

Initial speculation of the model. We'll include all the predictor variables and include all the interactive terms with colors at first.
```{r}
wine_rec_initial <- recipe(
  good_wine ~., data = wine_train) %>%
  step_rm(good_wine_names) %>%
  step_rm(quality) %>%
  step_interact(terms = ~starts_with("c_"):color) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_zv(all_predictors())
```

```{r}
wine_wflow_initial <- workflow() %>%
  add_model(wine_spec) %>%
  add_recipe(wine_rec_initial)
```

```{r}
wine_fit_initial <- wine_wflow_initial %>%
  fit(data = wine_train)
kable(tidy(wine_fit_initial), digits = 3)
```

From the table we can see that there are many variables with p-values
greater than 0.05 under 95% CI, so we can drop some of those valuables
if we were to conduct to a hypothesis test.

#### Full Logistic Model

$log~odds(good~wine) = 290.153 + 0.371*fixed~acidity - 3.217*volatile + 0.256* residual~sugar - 11.862*chlorides - 0.011*total.sulfur.dioxide - 304.879*density + 3.63*sulphates + 0.766*alcohol + 0.015*total.sulfur.dioxide x colorwhite - 265.692*density x colorwhite - 1.534*sulphates x colorwhite - 0.511*alcohol x colorwhite + 259.744*colorwhite$

```{r}
wine_rec_full <- recipe(
  good_wine ~., data = wine_train) %>%
  step_rm(good_wine_names) %>%
  step_interact(terms = ~starts_with("c_"):color) %>%
  step_rm(quality) %>%
  step_rm(c_free.sulfur.dioxide_x_colorwhite) %>%
  step_rm(c_chlorides_x_colorwhite) %>%
  step_rm(c_citric.acid_x_colorwhite) %>%
  step_rm(c_residual.sugar_x_colorwhite) %>%
  step_rm(c_volatile.acidity_x_colorwhite) %>%
  step_rm(c_fixed.acidity_x_colorwhite) %>%
  step_rm(c_free.sulfur.dioxide) %>%
  step_rm(c_citric.acid) %>%
  step_rm(c_pH) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_zv(all_predictors())
```

```{r}
wine_wflow_full <- workflow() %>%
  add_model(wine_spec) %>%
  add_recipe(wine_rec_full)
```

```{r}
wine_fit_full <- wine_wflow_full %>%
  fit(data = wine_train)
kable(tidy(wine_fit_full), digits = 3)
```

All the p-values are very small. We'll now apply this recipe to test set
to make prediction.

```{r}
wine_full_test <- wine_wflow_full %>%
  fit(data = wine_test)

tidy(wine_full_test) %>% 
  kable(digits = 3)
```

```{r}
wine_full_test_pred <- predict(wine_fit_full, wine_test, type = "prob") %>% 
  bind_cols(wine_test) 
wine_full_test_pred
```

#### Stepwise
```{r}
AIC_fit <- logistic_reg() %>%
  set_engine("glm") %>%
  fit(good_wine~.-c_citric.acid-quality-good_wine_names, data = wine_train)

AIC_fit <- repair_call(AIC_fit, data = wine_train)
AIC_fit_engine <- AIC_fit %>% extract_fit_engine() 
```

```{r}
best_AIC_model <- stepAIC(AIC_fit_engine, direction="forward", trace=FALSE)
tidy(best_AIC_model)
```

```{r}
best_AIC_pred <- predict(best_AIC_model, wine_test) %>% 
  bind_cols(wine_test)
best_AIC_pred <- best_AIC_pred %>% mutate(.pred_1 = exp(...1) / (1 + exp(...1))) %>% mutate(.pred_0 = 1 - .pred_1)
best_AIC_pred
```
#### Model Selection

```{r}
wine_full_test_pred %>% roc_curve(truth = as.factor(good_wine), .pred_0) %>% mutate(model = "Model 1") %>%
  bind_rows(best_AIC_pred %>% roc_curve(truth = as.factor(good_wine), .pred_0) %>% mutate(model = "Model 2")) %>%
  ggplot(aes(x = 1 - specificity, y = sensitivity, color = model)) +
  geom_line()
```

```{r}
wine_full_test_pred %>%
  roc_auc(truth = as.factor(good_wine), .pred_0)

best_AIC_pred %>%
  roc_auc(truth = as.factor(good_wine), .pred_0)
```

### Results

-   model interpretation
  - full model
  - best AIC model
  - AIC BIC ROC
-   model selection: do not include the interactive terms due to parsimonious. because the ROC does not improve much. 

### Discussion & Conclusion

-   what chemical components contribute to wine quality
-   future research suggestions
-   suggestions to wine valley?

### Reference

-   dataset
-   articles
