---
title: "What Makes a Good Glass of Wine?"
subtitle: "STA 210 - Summer 2022"
author: "Team 2 - Alicia Gong, Ashley Chen, Abdel Shehata, Claire Tan"
date: "6-16-2022"
format: pdf
editor: visual
editor_options: 
  markdown: 
    wrap: 72
---

```{r, echo=FALSE, message=FALSE}
#| label: load-pkg-data
#| message: false
#| warning: false
library(tidyverse)
library(tidymodels)
library(dplyr)
library(ggplot2)
library(cowplot)
library(knitr)
library(recipes)
library(caret)
library(InformationValue)
library(ISLR)
library(MASS)
library(nnet)
library(Stat2Data)
```

```{r echo=FALSE, message=FALSE}
#| label: load data
redwine <- read.csv("data/winequality-red.csv", sep = ";")
whitewine <- read.csv("data/winequality-white.csv", sep = ";")
redwine <- redwine %>% mutate(color="red")
whitewine <- whitewine %>% mutate(color="white")
wine <- redwine %>% full_join(whitewine)
wine <- slice(wine, sample(1:n()))
```

## Introduction and Data

#### Introduction

About 234 million hectoliters of wine were consumed in 2020, worldwide,
with the US making up approximately 14% of that consumption (Karlsson
2020). Since Wine composition and wine quality varies widely, it raises
the question: what makes a good wine?

\par

To answer that question, we will analyze the wine quality dataset from
Vinho Verde vineyard in Portugal, and more importantly try to narrow
down our question to make it possible for it to be supported by
evidence. Below is the introduction to our research:

\par

Project Goal: To identify variables that are important in explaining
variation in the response. "Vinho Verde" is the kind of wine exclusively
produced in the demarcated region of Vinho Verde in northwestern
Portugal. The Vinho Verde wine has its own unique system of production
and is only produced from the indigenous grape varieties of the region.
Vinho Verde region is one of the largest and oldest wine regions in the
world, and is home to thousands of producers, generating a wealth of
economic activity and jobs, and strongly contributes to the development
of Vinho province and the country. The Vinho Verde wine also enjoys high
reputation worldwide. It is recurrently awarded in national and
international competitions. The goal of this dataset is to model wine
quality based on physicochemical tests. We believe that this dataset can
also be used to analyze the relationship between different chemical
compositions and the ratings of wine quality. We believe that this
dataset can also be used to analyze what chemical factors are
attributable to the final rating of Vinho Verde wine. Our research may
shed light on future research and development directions for improving
the quality of Vinho Verde wine, which may also contribute to the
competitiveness of Portuguese wine industry.

\par

Our goal is to produce a classification model that best explains how
different chemical compositions of the Portuguese "Vinho Verde" wine
affects the variation of the wine quality.

#### Data Introduction

The Wine Quality dataset was collected from Vinho Verde wine Samples,
from the North of Portugal. The data was originally donated in 2009 by
Professor Cortez. The specific mechanism of the collection of the data
was lab work done on different wines to measure their chemical
attributes (like acidity etc.). The quality of the wine however was
obtained through the average rating of three wine experts. The dataset
is divided into two: Red wine and White wine. Red Wine has 1599
observations, and white wine has 4898 observations (each observation
being a specific wine). Information about the wine include but are not
limited to: pH, Density, Acidity, and alcohol content.

\par

Each observation is a specific wine from the Vinho Verde region. Thus,
there might be a little uncertainty in collecting the exact numbers for
each numbers. However, since the Vinho Verde region is a vast region
spreading 15500 hectares of vineyards in far-north Portugal, this
uncertainty shouldn't be significant in our analysis or project. Thus,
we will assume that the data are independent and random.

```{r datawrangling, echo=FALSE}
wine <- wine %>%
  mutate(good_wine = if_else(quality >= 7,"1","0"))
wine <- wine %>%
  mutate(good_wine = as.factor(good_wine))
wine <- wine %>%
  mutate(good_wine_names = if_else(good_wine=="1","Good wine","Bad or subpar wine"))

no1 <- colnames(wine)[1:11]
colnames(wine)[1:11] = paste("c_", no1, sep = "")
```

#### Exploratory Data Analysis

```{r distributions, echo=FALSE, message=FALSE, fig.align = "center"}
p1  <- ggplot(data = wine, aes(x = quality) ) + 
  geom_bar(fill = "pink") + 
  labs(x = "quality")

p2  <- ggplot(data = wine, aes(x = c_fixed.acidity) ) + 
  geom_histogram(fill = "pink") + 
  labs(x = "fixed acidity")

p3  <- ggplot(data = wine, aes(x = c_volatile.acidity) ) + 
  theme(axis.text=element_text(size=9)) + 
  geom_histogram(fill = "pink") + 
  labs(x = "volatile acidity")

p4  <- ggplot(data = wine, aes(x = c_citric.acid) ) + 
  theme(axis.text = element_text(size=9)) + 
  geom_histogram(fill = "pink") + 
  labs(x = "citric acid")

p5  <- ggplot(data = wine, aes(x = c_residual.sugar) ) +
  geom_histogram(fill = "pink") + 
  labs(x = "residual sugar")

p6  <- ggplot(data = wine, aes(x = c_chlorides) ) + 
  theme(axis.text = element_text(size = 11)) + 
  geom_histogram(fill = "pink") + 
  labs(x = "chlorides")

p7  <- ggplot(data = wine, aes(x = c_free.sulfur.dioxide) ) + 
  theme(axis.text = element_text(size=9)) + 
  geom_histogram(fill = "pink") + 
  labs(x = "free sulfur dioxide")

p8  <- ggplot(data = wine, aes(x = c_total.sulfur.dioxide) ) + 
  theme(axis.text = element_text(size=9)) + 
  geom_histogram(fill = "pink") + 
  labs(x = "total sulfur dioxide")

p9  <- ggplot(data = wine, aes(x = c_density) ) + 
  theme(axis.text = element_text(size = 7.5)) + 
  geom_histogram(fill= "pink") + 
  labs(x = "density")

p10 <- ggplot(data = wine, aes(x = c_pH) ) + 
  geom_histogram(fill = "pink") + 
  labs(x = "pH")

p11 <- ggplot(data = wine, aes(x = c_sulphates) ) + 
  theme(axis.text = element_text(size=9)) + 
  geom_histogram(fill= "pink") + 
  labs(x = "sulphates")

p12 <- ggplot(data = wine, aes(x = c_alcohol) ) + 
  geom_histogram(fill= "pink") + 
  labs(x = "alcohol")

plot_grid(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, p11, p12, ncol = 3, nrow = 4)
```

```{r, echo=FALSE, fig.height=7, fig.width=14, fig.align='center'}
a1 <- ggplot(wine, aes(y = c_alcohol, x = as.factor(good_wine), color = color)) + 
  geom_boxplot() + 
  labs(x = "good wine", y = "alcohol")

a2 <- ggplot(wine, aes(y = c_density, x = as.factor(good_wine), color = color)) + 
  geom_boxplot() + 
  labs(x = "good wine", y = "density")

a3 <- ggplot(wine, aes(y = c_volatile.acidity, x = as.factor(good_wine), color = color)) + 
  geom_boxplot() + 
  labs(x = "good wine", y = "volatile acidity")

a4 <- ggplot(wine, aes(y = c_chlorides, x = as.factor(good_wine), color = color)) + 
  geom_boxplot() + 
  labs(x = "good wine", y = "chlorides")

a5 <- ggplot(wine, aes(y = c_sulphates, x = as.factor(good_wine), color = color)) + 
  geom_boxplot() + 
  labs(x = "good wine", y = "sulphates")

a6 <- ggplot(wine, aes(y = c_pH, x = as.factor(good_wine), color = color)) + 
  geom_boxplot() + 
  labs(x = "good wine", y = "pH")

a7 <- ggplot(wine, aes(y = c_total.sulfur.dioxide, x = as.factor(good_wine), color = color)) + 
  geom_boxplot() + 
  labs(x = "good wine", y = "Total Sulfur Dioxide")

a8 <- ggplot(wine, aes(y = c_residual.sugar, x = as.factor(good_wine), color = color)) + 
  geom_boxplot() + 
  labs(x = "good wine", y = "Residual Sugar")

a9 <- ggplot(wine, aes(y = c_free.sulfur.dioxide, x = as.factor(good_wine), color = color)) + 
  geom_boxplot() + 
  labs(x = "good wine", y = "Free Sulfur Dioxide")

a10 <- ggplot(wine, aes(y = c_fixed.acidity, x = as.factor(good_wine), color = color)) + 
  geom_boxplot() + 
  labs(x = "good wine", y = "Fixed Acidity")

a11 <- ggplot(wine, aes(y = c_citric.acid, x = as.factor(good_wine), color = color)) + 
  geom_boxplot() + 
  labs(x = "good wine", y = "Citric Acid")

a12 <- ggplot(wine, aes(y = quality, x = as.factor(good_wine), color = color)) + 
  geom_boxplot() + 
  labs(x = "good wine", y = "Quality")

plot_grid(a1, a2, a3, a4, a5, a6, a7, a8, a9, a10, a11, a12, ncol = 4, nrow = 3)
```

## Methodology

#### Best AIC Model

Given that there are many predictor variables in the model, we dicide to conduct a stepwise AIC test to help us shave off the unrelated predictors. This is our preliminary screening of the variables, so we did not consider interactive effects here. 

We can operationalize the best AIC model as the model with the
lowest AIC value to supplement which variables to add to or omit from to
choose the model with the smallest amount of error or lowest residual
sum of squares. AIC is an estimator of in-sample prediction error and a
lower AIC values can indicate a more parsimonious model.

In this stepwise AIC test, an algorithm uses specific procedures in
which the AIC values of different models calculated to determine what
covariates are added to or removed from the model and this process is
repeated several times in both directions. After conducting the stepwise
AIC test, we found that the combination of variables with the lowest AIC
values were fixed acidity, volatile acidity, residual sugar, chlorides,
free sulfur dioxide, total sulfur dioxide, density, pH, and sulphates.

We split the data into 25% testing set and 75% training set. The training set is used to develop models, and the test set is used for estimating a final, unbiased assessment of the model’s performance.

```{r datasplit, echo=FALSE}
set.seed(222)
wine_split <- initial_split(wine, prop = 3/4)
wine_train <- training(wine_split)
wine_test <- training(wine_split)
```

```{r, echo=FALSE}
AIC_fit <- logistic_reg() %>%
  set_engine("glm") %>%
  fit(good_wine~.-quality-good_wine_names, data = wine_train)

AIC_fit <- repair_call(AIC_fit, data = wine_train)
AIC_fit_engine <- AIC_fit %>% extract_fit_engine() 
```

#### Model output

```{r, echo=FALSE}
best_AIC_model <- stepAIC(AIC_fit_engine, direction="forward", trace=FALSE)
kable(tidy(best_AIC_model), digits = 3)
```

$logodds(good~wine) = 337.48 + 0.533 \times fixed~acidity - 3.508 \times volatile~acidity - 0.453 \times citric acid + 0.198 \times residual~sugar - 8.59 \times chlorides + 0.009 \times free~sulfur~dioxide - 0.003 \times total~sulfur~dioxide - 359.511 \times density + 2.707 \times pH + 2.294 \times sulphates + 0.545 \times alcohol - 0.52 \times colorwhite$

After conducting the stepwise AIC test, the following variables are selected:

- Fixed Acidity: the amount of acid in wine that’s not volatile (do not evaporate fast).

- Volatile Acidity: the amount of acetic acid in wine.

- Citric Acid: citric acid is found in small quantities, and can add
freshness and flavor to wines.

- Residual Sugar: the amount of sugar left after fermatation stops.

- Chlorides: the amount of salt in the wine.

- Free Sulfur Dioxide: the free amount of $SO_2$ exists in equilibrium between molecular $SO_2$ and bisulfite ion. 

- Total Sulfur Dioxide: the total amount of free and bounded $SO_2$.

- Density: the density of the liquid. 

- pH: hydrogen ion concentration

- Sulphates: produced by yeast, protecting wine against oxidation.

- Alcohol: the percentage of wine's alcohol content.


```{r, include=FALSE}
best_AIC_fit <- predict(best_AIC_model, wine_train) %>% bind_cols(wine_train)
best_AIC_fit <- best_AIC_fit %>% mutate(.pred_1 = exp(...1) / (1 + exp(...1))) %>% mutate(.pred_0 = 1 - .pred_1)
```

```{r, echo=FALSE, fig.width=3, fig.height=2, fig.align='center'}
best_AIC_pred <- predict(best_AIC_model, wine_test) %>% bind_cols(wine_test)
best_AIC_pred <- best_AIC_fit %>% mutate(.pred_1 = exp(...1) / (1 + exp(...1))) %>% mutate(.pred_0 = 1 - .pred_1)
best_AIC_pred %>% roc_curve(truth = as.factor(good_wine), .pred_0) %>%
  ggplot(aes(x = 1 - specificity, y = sensitivity)) +
  geom_line() +
  labs(
    title = "Best AIC Model's ROC Curve on Prediction"
  )
```

## Model Evaluation

#### Full Model

Then, considering that we have two types of wine--red wine and white wine, and we have seen potential interactive effects between the color of wine and the level of chemical components from the EDA, we decided to add interactive terms between color and othe predictor variables.
We compare the best AIC model with another logistic model with
interactive terms. We'll call the best AIC model as the reduced model,
and the logistic model with interactive terms as the full model.

```{r, echo=FALSE}
wine_spec <- logistic_reg() %>% set_engine("glm")
```

```{r, echo=FALSE}
wine_rec_full <- recipe(
  good_wine ~., data = wine_train) %>%
  step_rm(good_wine_names) %>%
  step_rm(quality) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_interact(terms = ~starts_with("c_"):starts_with("color")) %>%
  step_rm(c_alcohol_x_color_white) %>%
  step_zv(all_predictors())
```

```{r, echo=FALSE}
wine_wflow_full <- workflow() %>%
  add_model(wine_spec) %>%
  add_recipe(wine_rec_full)

wine_fit_full <- wine_wflow_full %>%
  fit(data = wine_train)
kable(tidy(wine_fit_full), digits = 3)
```

$logodds(good~wine) = 377.14 + 0.419 \times fixed~acidity - 1.922 \times volatile~acidity - 0.854 \times citric~acid + 0.31 \times residual~sugar - 9.443 \times chlorides + 0.018 \times free.sulfur.dioxide - 0.022 \times total.sulfur.dioxide - 394.91 \times density + 1.465 \times pH + 3.924 \times sulphates + 0.422 \times alcohol + 53.152 \times colorwhite - 0.054 fixed~acidity:colorwhite - 2.545 volatile~acidity :colorwhite - 1.853 citric~acid:colorwhite - 0.089 residual~sugar:colorwhite - 2.774 chlorides:colorwhite - 0.008 free~sulfur~dioxide:colorwhite + 0.021 total~sulfur~dioxide:colorwhite$

In this model, red wine is picked as the baseline. For example, the odds of being classified as 'good wine' for white wine are expected to be exp(-0.054) = 0.947 times the odds for red wine, holding all else constant.

From the EDA, we can see that there is no significant difference between the alcohol concentration between red and white wine, so we remove the interactive terms of alcohol and colorwhite.

```{r, echo=FALSE}
wine_full_test <- wine_wflow_full %>%
  fit(data = wine_test)

wine_full_fit <- predict(wine_fit_full, wine_train, type = "prob") %>% bind_cols(wine_train)
```

## Model Selection

We fit both the full model with interactive terms and the initial Best AIC model without interactive terms to the training data.

```{r, echo=FALSE, fig.width=6, fig.height=5, fig.align='center'}
wine_full_fit %>% roc_curve(truth = as.factor(good_wine), .pred_0) %>% mutate(model = "Model 1") %>%
  bind_rows(best_AIC_fit %>% roc_curve(truth = as.factor(good_wine), .pred_0) %>% mutate(model = "Model 2")) %>%
  ggplot(aes(x = 1 - specificity, y = sensitivity, color = model)) +
  geom_line() +
  labs(
    title = "Two Models' ROC Curves"
  )
```

```{r, echo=FALSE}
aoc_logistic <- wine_full_fit %>% 
  roc_auc(truth = as.factor(good_wine), .pred_0) %>%
  mutate(model = "Full Model")

aoc_AIC <- best_AIC_fit %>% 
  roc_auc(truth = as.factor(good_wine), .pred_0) %>%
  mutate(model = "Best AIC Model")

twoaoc <- bind_rows(aoc_logistic, aoc_AIC) %>%
arrange(.metric) %>% relocate(model, .before = .metric)

kable(twoaoc, digits = 3)
```

The two models' ROC curves and ROC-AUC are very close, suggesting that
the two models have similar performances.

#### Cross Validation

We also conduct a cross validation to further assess the two models'
performances. To perform the cross validation, we split the training
data into 10 folds. From the plots we can see that the two models have
similar accuracy and the ROC-AUC for all 10 folds.

```{r, echo=FALSE}
best_AIC_rec <- recipe(
  good_wine ~ c_fixed.acidity + c_volatile.acidity + 
    c_citric.acid + c_residual.sugar + c_chlorides + c_free.sulfur.dioxide + 
    c_total.sulfur.dioxide + c_density + c_pH + c_sulphates + 
    c_alcohol + quality + color + good_wine_names, data = wine_train) %>%
  step_rm(quality) %>%
  step_rm(good_wine_names) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_zv(all_predictors())

wine_wflow_reduce <- workflow() %>%
  add_model(wine_spec) %>%
  add_recipe(best_AIC_rec)
```

```{r, echo=FALSE}
set.seed(222)
folds <- vfold_cv(wine_train, v = 10)
```

```{r, echo=FALSE}
wine_full_rs <- wine_wflow_full %>%
  fit_resamples(folds)

wine_AIC_rs <- wine_wflow_reduce %>%
  fit_resamples(folds)
```

```{r, echo=FALSE}
metrics_full <- collect_metrics(wine_full_rs, summarize = FALSE) %>% mutate(model = "Model 1 - Full")
metrics_reduce <- collect_metrics(wine_AIC_rs, summarize = FALSE) %>% mutate(model = "Model 2 - Reduce")
```

```{r, echo=FALSE, fig.height=4, fig.width=8, fig.align='center'}
metrics <- bind_rows(metrics_full, metrics_reduce) %>%
arrange(.metric)
acc_p <- ggplot(metrics %>% filter(.metric == "accuracy"), 
                aes(x = id, y = .estimate,
                group = model, color = model)) + geom_point() + geom_line() + labs(title = "Accuracy", x = "Folds", y = "Estimate") + theme_bw()

aoc_p <- ggplot(metrics %>% filter(.metric == "roc_auc"), aes(x = id, y = .estimate,
    group = model, color = model)) + geom_point() + geom_line() + labs(title = "ROC-AUC", x = "Folds", y = "Estimate") + theme_bw()

plot_grid(acc_p, aoc_p, nrow = 2)
```

## Limitations

```{r, include=FALSE}
emplogitplot1(good_wine ~ c_fixed.acidity, 
              data = wine, 
              ngroups = 20)

emplogitplot1(good_wine ~ c_volatile.acidity, 
              data = wine, 
              ngroups = 20)

emplogitplot1(good_wine ~ c_free.sulfur.dioxide, 
              data = wine, 
              ngroups = 30)

emplogitplot1(good_wine ~ c_residual.sugar, 
              data = wine, 
              ngroups = 30)

emplogitplot1(good_wine ~ c_chlorides, 
              data = wine, 
              ngroups = 30)

emplogitplot1(good_wine ~ c_total.sulfur.dioxide, 
              data = wine, 
              ngroups = 30)

emplogitplot1(good_wine ~ c_density, 
              data = wine, 
              ngroups = 30)

emplogitplot1(good_wine ~ c_pH + color, 
              data = wine, 
              ngroups = 30)

emplogitplot1(good_wine ~ c_sulphates, 
              data = wine, 
              ngroups = 30)

emplogitplot1(good_wine ~ c_alcohol + color, 
              data = wine, 
              ngroups = 20)
```

When conducting a logistic regression, it is important to check the
conditions are satisfied. When evaluating if the log-odds have a linear
relationship with the predictors, we find that the conditions for
linearity are not satisfied for all the variables. The linearity
conditions are particularly not fulfilled for the variables free sulfur
dioxide (the spread takes on a curved shape), total sulfur dioxide (the
spread has similarly curved shape), and sulphates (the data points
appear to be randomly dispersed).

The independence conditions, however, are satisfied. The Vinho Verde
region is a vast region spreading 15500 hectares of vineyards in
far-north Portugal allowing the observations to be collected
independently from each other and the accumulated data are not collected
across an extended period of time.

## Results

We first conducted a pairwise test to obtain a Best AIC Model with no
interactive terms, which is our reduced model. Then we include the
interactive terms and build another logistic model, as our full model.
We compared these two models' ROC curves, AIC, BIC. We also check the
two models' accuracy and ROC-AUC through cross-validation. We found out
that the two models have similar performances when predicting the
quality of wine. Due to the principles of parsimonious, we prefer the
simpler model. Therefore, we choose the reduced Best AIC Model as our
best model.

$logodds(good~wine) = 337.48 + 0.533 \times fixed~acidity - 3.508 \times volatile~acidity - 0.453 \times citric acid + 0.198 \times residual~sugar - 8.59 \times chlorides + 0.009 \times free~sulfur~dioxide - 0.003 \times total~sulfur~dioxide - 359.511 \times density + 2.707 \times pH + 2.294 \times sulphates + 0.545 \times alcohol - 0.52 \times colorwhite$

## Discussion & Conclusion

From our analysis, we found that the variables that contribute
significantly to wine quality were fixed acidity, volatile acidity,
residual sugar, chlorides, free sulfur dioxide, total sulfur dioxide,
density, pH, and sulfates. Variables with biggest coefficients are
chlorides, volatile acidity, pH and sulphates. It is worth-noticing that
although the coefficient of density is very high, wine's density varies
between approximate 0.99 to 1, therefore it is impossible for wine'
density to jump 1 unit. So we do not include density as a influential
factor.

-   Negative effects:

    Chlorides: a major contributor to saltiness.

    Volatile Acidity: a kind of acid which produces vinegar-like taste
    in wine.

-   Positive effects:

    pH: low pH wines will taste tart and crisp.

    Sulphates: sulphates protect wine against oxidation, which can affect the color and taste of wine.

In the future, to improve the accuracy of the model, different stepwise
algorithms (p-values for example) can be evaluated and adjusted. There
are many different methods such as through feature engineering, viewing
potential interaction terms, or other performance measurements such as
machine learning algorithms to better predict our results.

## Reference

Coli, Marina Sonegheti et al. Chloride concentration in red wines:
influence of terroir and grape type. Food Science and Technology
(Campinas) \[online\]. 2015, v. 35, n. 1 \[Accessed 17 June 2022\] , pp.
95-99. Available from: <https://doi.org/10.1590/1678-457X.6493>. ISSN
1678-457X. https://doi.org/10.1590/1678-457X.6493.

Paulo Cortez, University of Minho, Guimarães, Portugal,
<http://www3.dsi.uminho.pt/pcortez> A. Cerdeira, F. Almeida, T. Matos
and J. Reis, Viticulture Commission of the Vinho Verde Region(CVRVV),
Porto, Portugal, 2009.
