---
title: "What Makes a Good Glass of Wine?"
subtitle: "STA 210 - Summer 2022"
author: "Team 2 - Alicia Gong, Ashley Chen, Abdel Shehata, Claire Tan"
date: "6-16-2022"
format: pdf
editor: visual
editor_options: 
  markdown: 
    wrap: 72
---

```{r, echo=FALSE, message=FALSE}
#| label: load-pkg-data
#| message: false
#| warning: false
library(tidyverse)
library(tidymodels)
library(dplyr)
library(ggplot2)
library(cowplot)
library(knitr)
library(recipes)
library(caret)
library(InformationValue)
library(ISLR)
library(MASS)
library(nnet)
library(Stat2Data)
```

```{r echo=FALSE, message=FALSE}
#| label: load data
set.seed(222)
redwine <- read.csv("data/winequality-red.csv", sep = ";")
whitewine <- read.csv("data/winequality-white.csv", sep = ";")
redwine <- redwine %>% mutate(color="red")
whitewine <- whitewine %>% mutate(color="white")
wine <- redwine %>% full_join(whitewine)
wine <- slice(wine, sample(1:n()))
```

### Introduction and Data

#### Introduction

In 2020 alone, about 234 million hectoliters of wine were consumed
worldwide, with the US making up approximately 14% of that consumption
(Karlsson 2020). Both wine consumption and wine production are essential
parts of many cultures and economies around the world. For some
communities, wine production is their primary economic means to
survival, and some wine regions are even recognized by UNESCO as World
Hertiage sites. Thus, it is important to be able to assess the quality
of wine based on its physicochemical components. Since wine composition
and wine quality varies widely, it raises the question: what makes a
good wine? More specifically, what chemical compositions of wine affect
the variation of wine quality? To answer this question, we will analyze
the wine quality dataset from Vinho Verde vineyard in Portugal, and more
importantly try to narrow down our question to make it possible for it
to be supported by evidence.

\par

"Vinho Verde" wine is exclusively produced in the demarcated region of
Vinho Verde in northwestern Portugal, which is one of the largest and
oldest wine regions in the world and home to thousands of producers,
generating a wealth of economic activity and jobs and strongly
contributing to the development of the Vinho province and the country as
a whole. This wine has its own unique system of production, as it is
only produced from the indigenous grape varieties of the region.
Additionally, the wine produced in Vinho Verde also enjoys high
reputation worldwide, and is recurrently awarded in national and
international competitions.

\par

We are curious about what exactly makes a good glass of Vinho Verde
wine, from a scientific standpoint. Wine-drinkers commonly relate the
quality of a wine to its aroma or sweetness, but these standards are
subjective to the person drinking the wine. Because the chemical factors
of wine are what contribute to these varying characteristics, we predict
that chemical components which contribute to wine's aroma, sweetness,
ripeness and freshness will be influential to wine's quality measure
points.

\par

By investigating and analyzing the chemical factors attributable to the
final rating of Vinho Verde wine, our research may shed light on future
research and development directions for improving the quality of Vinho
Verde wine, which may also contribute to the competitiveness of
Portuguese wine industry. Thus, the overall goal of this project is to
produce a classification model that best explains how different chemical
compositions of the Portuguese "Vinho Verde" wine affects the variation
of the wine quality.

#### Data Introduction

The Wine Quality dataset was collected from Vinho Verde wine Samples,
from northern Portugal. The data was originally donated in 2009 by
Professor Cortez. The specific mechanism of the collection of the data
was lab work done on different wines to measure their chemical
attributes (such as acidity). The quality of the wine was obtained
through the average rating of three wine experts. The dataset is divided
into two: Red wine and White wine. Red Wine has 1599 observations, and
white wine has 4898 observations (each observation being a specific
wine). Key variables include:

-   `quality`: wine quality, rated on a scale from 0-9.
-   `volatile.acidity`: amount of gaseous acid in wine
-   `chlorides`: amount of salt in wine
-   `total.sulfur.dioxide`: amount of free and bound forms of $SO_2$ in
    wine
-   `pH`: acidity of wine
-   `sulphates`: amount of sulphates in wine
-   `alcohol`: percent alcohol content of the wine

\par

In our analysis, we split `quality` into "good wine" and "bad wine",
defining "good wine" as wine that receives a quality score higher than
7. We based this on the system of the wine magazine 'Wine Spectator', in
which a score of at least '80' (on a 100-point scale) is defined as
'Good'. In our data, the highest score a wine can get is 9, and 0.8 \* 9
= 7.2, so we chose 7 as our cut-off point, and created a new variable
`good_wine` that is 1 for good wine, and 0 otherwise,

```{r datawrangling, echo=FALSE}
set.seed(222)
wine <- wine %>%
  mutate(good_wine = if_else(quality >= 7,"1","0"))
wine <- wine %>%
  mutate(good_wine = as.factor(good_wine))
wine <- wine %>%
  mutate(good_wine_names = if_else(good_wine=="1","Good wine","Bad or subpar wine"))

no1 <- colnames(wine)[1:11]
colnames(wine)[1:11] = paste("c_", no1, sep = "")
```

#### Exploratory Data Analysis

We first visualized the distributions of potential predictor variables
through histograms. Then, we visualized the relationships between key
predictor variables and the response variable `good_wine` through
boxplots, and also investigated possible interaction effects between
these predictor variables and the color of the wine.

```{r distributions, echo=FALSE, message=FALSE, fig.align = "center"}
p1  <- ggplot(data = wine, aes(x = good_wine) ) + 
  geom_bar(fill = "pink") + 
  labs(x = "quality")

p2  <- ggplot(data = wine, aes(x = c_fixed.acidity) ) + 
  geom_histogram(fill = "pink") + 
  labs(x = "fixed acidity")

p3  <- ggplot(data = wine, aes(x = c_volatile.acidity) ) + 
  theme(axis.text=element_text(size=9)) + 
  geom_histogram(fill = "pink") + 
  labs(x = "volatile acidity")

p4  <- ggplot(data = wine, aes(x = c_citric.acid) ) + 
  theme(axis.text = element_text(size=9)) + 
  geom_histogram(fill = "pink") + 
  labs(x = "citric acid")

p5  <- ggplot(data = wine, aes(x = c_residual.sugar) ) +
  geom_histogram(fill = "pink") + 
  labs(x = "residual sugar")

p6  <- ggplot(data = wine, aes(x = c_chlorides) ) + 
  theme(axis.text = element_text(size = 11)) + 
  geom_histogram(fill = "pink") + 
  labs(x = "chlorides")

p7  <- ggplot(data = wine, aes(x = c_free.sulfur.dioxide) ) + 
  theme(axis.text = element_text(size=9)) + 
  geom_histogram(fill = "pink") + 
  labs(x = "free sulfur dioxide")

p8  <- ggplot(data = wine, aes(x = c_total.sulfur.dioxide) ) + 
  theme(axis.text = element_text(size=9)) + 
  geom_histogram(fill = "pink") + 
  labs(x = "total sulfur dioxide")

p9  <- ggplot(data = wine, aes(x = c_density) ) + 
  theme(axis.text = element_text(size = 7.5)) + 
  geom_histogram(fill= "pink") + 
  labs(x = "density")

p10 <- ggplot(data = wine, aes(x = c_pH) ) + 
  geom_histogram(fill = "pink") + 
  labs(x = "pH")

p11 <- ggplot(data = wine, aes(x = c_sulphates) ) + 
  theme(axis.text = element_text(size=9)) + 
  geom_histogram(fill= "pink") + 
  labs(x = "sulphates")

p12 <- ggplot(data = wine, aes(x = c_alcohol) ) + 
  geom_histogram(fill= "pink") + 
  labs(x = "alcohol")

plot_grid(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, p11, p12, ncol = 3, nrow = 4)
```

```{r, echo=FALSE, fig.height=4, fig.width=8, fig.align='center'}
a1 <- ggplot(wine, aes(y = c_alcohol, x = as.factor(good_wine), color = color)) + 
  geom_boxplot() + 
  labs(x = "good wine", y = "alcohol")

a2 <- ggplot(wine, aes(y = c_density, x = as.factor(good_wine), color = color)) + 
  geom_boxplot() + 
  labs(x = "good wine", y = "density")

a3 <- ggplot(wine, aes(y = c_volatile.acidity, x = as.factor(good_wine), color = color)) + 
  geom_boxplot() + 
  labs(x = "good wine", y = "volatile acidity")

a4 <- ggplot(wine, aes(y = c_chlorides, x = as.factor(good_wine), color = color)) + 
  geom_boxplot() + 
  labs(x = "good wine", y = "chlorides")

a5 <- ggplot(wine, aes(y = c_sulphates, x = as.factor(good_wine), color = color)) + 
  geom_boxplot() + 
  labs(x = "good wine", y = "sulphates")

a6 <- ggplot(wine, aes(y = c_pH, x = as.factor(good_wine), color = color)) + 
  geom_boxplot() + 
  labs(x = "good wine", y = "pH")

a7 <- ggplot(wine, aes(y = c_total.sulfur.dioxide, x = as.factor(good_wine), color = color)) + 
  geom_boxplot() + 
  labs(x = "good wine", y = "Total Sulfur Dioxide")

a8 <- ggplot(wine, aes(y = c_residual.sugar, x = as.factor(good_wine), color = color)) + 
  geom_boxplot() + 
  labs(x = "good wine", y = "Residual Sugar")

a9 <- ggplot(wine, aes(y = c_free.sulfur.dioxide, x = as.factor(good_wine), color = color)) + 
  geom_boxplot() + 
  labs(x = "good wine", y = "Free Sulfur Dioxide")

plot_grid(a1, a2, a3, a4, a5, a6, a7, a8, a9, ncol = 3, nrow = 3)
```

From this, we can see that most variables are normally distributed, but
residual sugar, chlorides, and sulphates are slightly skewed to the
right, while free sulfur dioxide and alcohol have a strongly
right-skewed distribution. Additionally, the distribution of good and
bad wine is quite unbalanced, as the number of observations categorized
as not good greatly outnumbers the amount categorized as good. From the
boxplot, we can see that for almost all variables expect alcohol, there
exists a significant difference between red and white wine. Therefore,
we will consider adding the interaction between color and other
predictor variables in our model.

```{r, include=FALSE}
pairs(quality ~ c_fixed.acidity + c_volatile.acidity + c_citric.acid + c_density + c_residual.sugar + c_chlorides, data=wine)
pairs(quality ~ c_free.sulfur.dioxide + c_total.sulfur.dioxide + c_pH + c_sulphates + c_alcohol, data=wine)
```

```{r, echo=FALSE}
# correlation coefficients for potential multicollinearity
c1 <- cor(wine$c_fixed.acidity, wine$c_citric.acid)
c2 <- cor(wine$c_fixed.acidity, wine$c_density)
c3 <- cor(wine$c_volatile.acidity, wine$c_citric.acid)
c4 <- cor(wine$c_density, wine$c_residual.sugar)
c5 <- cor(wine$c_free.sulfur.dioxide, wine$c_total.sulfur.dioxide)

Variables <- c("fixed acidity & citric acid",
                            "fixed acidity & density", 
                            "volatile acidity and citric acid", 
                            "density & residual sugar", 
                            "free sulfur dioxide and total sulfur dioxide")
R_Coefficients <- c(c1, c2, c3, c4, c5)

df <- data.frame(Variables, R_Coefficients)
kable(tibble(df), digits = 3)
```

Lastly, we calculated the correlation coefficients for some of our
predictor variables to investigate multicollinearity, which is the
ocurrence of high intercorrelations among two or more independent
variables in a multiple regression model. This is a problem because it
undermines the statistical significance of the predictor variables and
makes it difficult to establish relationships. From our calculations, we
can see that the correlation between free.sulfur.dioxide and
total.sulfur.dioxide is 0.72, indicating a strong positive linear
relationship, so multicollinearity exists between these two variables.

### Methodology

#### Best AIC Model

To select the most optimal model to fit our data, we decided to conduct
a step-wise AIC model. We can operationalise this as the model with the
lowest AIC value to supplement which variables to add to or omit from to
choose the model with the smallest amount of error or lowest residual
sum of squares. AIC is an estimator of in-sample prediction error and a
lower AIC values can indicate a more parsimonious model.

In this step-wise AIC test, an algorithm uses specific procedures in
which the AIC values of different models are calculated to determine
what covariates are added to or removed from the model, and this process
is repeated several times in both directions. In our EDA, we found
multicollinearity in the strong positive linear relationship between
free.sulfur.dioxide and total.sulfur.dioxide. We also know from the data
description that total.sulfur dioxide includes free.sulfur.dioxide -
whose concentration level could impact the smell and taste of the wine,
thus impacting quality. So it is sufficient to include only one of these
two variables in the model, and we choose free.sulfur.dioxide. After
conducting the step-wise AIC test, we found that the combination of
variables with the lowest AIC values were fixed acidity, volatile
acidity, residual sugar, chlorides, free sulfur dioxide, density, pH,
and sulphates.

We split the data into 25% testing set and 75% training set. The
training set is used to develop models, and the test set is used for
estimating a final, unbiased assessment of the model's performance.

```{r datasplit, echo=FALSE}
set.seed(1234)

wine_split <- initial_split(wine, prop = 3/4)
wine_train <- training(wine_split)
wine_test <- training(wine_split)
```

```{r, echo=FALSE}
set.seed(1234)
AIC_fit <- logistic_reg() %>%
  set_engine("glm") %>%
  fit(good_wine~.-c_total.sulfur.dioxide-quality-good_wine_names, data = wine_train)

AIC_fit <- repair_call(AIC_fit, data = wine_train)
AIC_fit_engine <- AIC_fit %>% extract_fit_engine() 
```

#### Model output

```{r, echo=FALSE}
best_AIC_model <- stepAIC(AIC_fit_engine, direction="forward", trace=FALSE)
kable(tidy(best_AIC_model), digits = 3)
```

In this model, red wine is picked as the baseline. For example, the odds
of being classified as 'good wine' for white wine are expected to be
exp(-0.826) = 0.30 times the odds for red wine, holding all else
constant.

```{r, include=FALSE}
set.seed(222)
best_AIC_fit <- predict(best_AIC_model, wine_train) %>% bind_cols(wine_train)
best_AIC_fit <- best_AIC_fit %>% mutate(.pred_1 = exp(...1) / (1 + exp(...1))) %>% mutate(.pred_0 = 1 - .pred_1)
```

After conducting the step- wise AIC test, the following variables are
selected:

-   `fixed.acidity`: the amount of acid in wine that's not volatile (do
    not evaporate fast)
-   `volatile.acidity`: the amount of acetic acid in wine
-   `citric.acid`: found in small quantities and can add freshness and
    flavor to wines
-   `residual.sugar`: amount of sugar left after fermentation
-   `chlorides`: amount of salt in wine
-   `free.sulfur.dioxide`: free amount of $SO_2$ exists in equilibrium
    between molecular $SO_2$ and bisulfite ion
-   `density`: density of wine measured in g/ml
-   `pH`: acidity of wine and hydrogen ion concentration
-   `sulphates`: produced by yeast, protecting wine against oxidation
-   `alcohol`: percent alcohol content of the wine

\pagebreak

#### Model Evaluation

#### Full Model

Then, considering that we have two types of wine--red wine and white
wine, and we have seen potential interactive effects between the color
of wine and the level of chemical components from the EDA, we decided to
add interactive terms between color and other predictor variables. We
compare the best AIC model with another logistic model that includes
interactive terms. We will call the best AIC model the reduced model and
the logistic model with interactive terms the full model.

```{r, echo=FALSE}
wine_spec <- logistic_reg() %>% set_engine("glm")
```

```{r, echo=FALSE}
set.seed(222)
wine_rec_full2 <- recipe(
  good_wine ~., data = wine_train) %>%
  step_rm(quality) %>%
  step_rm(good_wine_names) %>%
  step_rm(c_total.sulfur.dioxide) %>%
  step_rm(c_citric.acid) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_interact(terms = ~starts_with("c_"):starts_with("color")) %>%
  step_rm(c_alcohol_x_color_white) %>%
  step_rm(c_density_x_color_white) %>%
  step_rm(c_pH_x_color_white) %>%
  step_rm(c_residual.sugar_x_color_white) %>%
  step_rm(c_chlorides_x_color_white) %>%
  step_zv(all_predictors())
```

```{r, echo=FALSE}
set.seed(222)
wine_wflow_full2 <- workflow() %>%
  add_model(wine_spec) %>%
  add_recipe(wine_rec_full2)

wine_fit_full2 <- wine_wflow_full2 %>%
  fit(data = wine_train)
kable(tidy(wine_fit_full2), digits = 3)
```

$\log\Big(\frac{\hat\pi}{1-\hat\pi}\Big) = 426.077 + 0.574 \times fixed~acidity - 3.002 \times volatile~acidity + 0.222 \times residual~sugar - 8.897 \times chlorides + 0.017 \times free.sulfur.dioxide - 447.221 \times density + 2.303 \times pH + 3.395 \times sulphates + 0.436 \times alcohol + 1.277 \times colorwhite - 0.229 \times fixed~acidity:colorwhite - 1.216 \times volatile~acidity:colorwhite + 0.026 \times free~sulfur~dioxide:colorwhite - 1.310 \times sulphates:colorwhite$

The interactive terms between predictor variables and colorwhite are
used to assess the interactive effects between other predictor variables
and wine's color. Red wine is used as baseline here. From the EDA, we
can see that there is no significant difference between the alcohol
concentration between red and white wine, so we did not consider the
interactive effects between alcohol and colorwhite. We also removed
citric acid, the interactive effect term of density:color_white,
pH:color_white, residual.sugar:color_white, chlorides:color_white due to
their high p-values.

```{r, echo=FALSE}
wine_full_test <- wine_wflow_full2 %>%
  fit(data = wine_test)

wine_full_fit2 <- predict(wine_fit_full2, wine_train, type = "prob") %>% bind_cols(wine_train)
```

#### Model Selection

One measure to compare the two classification models above is to use the
area under the ROC curve to assess their performance and determine which
model performed more optimally. We fit both the full model with
interactive terms and the initial Best AIC model without interactive
terms to the training data.

```{r, echo=FALSE, fig.width=3, fig.height=2, style= "float:right;position: relative; top: -80px;"}
wine_full_fit2 %>% roc_curve(truth = as.factor(good_wine), .pred_0) %>% mutate(model = "Model 1") %>%
  bind_rows(best_AIC_fit %>% roc_curve(truth = as.factor(good_wine), .pred_0) %>% mutate(model = "Model 2")) %>%
  ggplot(aes(x = 1 - specificity, y = sensitivity, color = model)) +
  geom_line() + labs(title = "Two Models' ROC curves")
```

The two ROC curves are very close, suggesting that the two models'
performances are similar.

We also calculated the area under the ROC curve (AUC) to summarize the
performance of each classifier into one single measure. The AUC value
tells how well the model is capable of distinguishing between two binary
classes. The higher the AUC, the better the model is predicting good
wine quality and good wine quality and bad wine quality as bad wine
quality.

We found that the AUC value of the logistic model was 0.815 and the AUC
value of the best AIC model was 0.811, suggesting that the two models
have similar performances.

```{r, echo=FALSE}
aoc_logistic <- wine_full_fit2 %>% 
  roc_auc(truth = as.factor(good_wine), .pred_0) %>%
  mutate(model = "Logistic Model")

aoc_AIC <- best_AIC_fit %>% 
  roc_auc(truth = as.factor(good_wine), .pred_0) %>%
  mutate(model = "Best AIC Model")

twoaoc <- bind_rows(aoc_logistic, aoc_AIC) %>%
arrange(.metric) %>% relocate(model, .before = .metric)

kable(twoaoc, digits = 3)
```

```{r, include=FALSE}
best_AIC_rec <- recipe(
  good_wine ~ c_fixed.acidity + c_volatile.acidity + 
    c_citric.acid + c_residual.sugar + c_chlorides + c_free.sulfur.dioxide + 
    c_total.sulfur.dioxide + c_density + c_pH + c_sulphates + 
    c_alcohol + quality + color + good_wine_names, data = wine_train) %>%
  step_rm(c_citric.acid) %>%
  step_rm(c_total.sulfur.dioxide) %>%
  step_rm(quality) %>%
  step_rm(good_wine_names) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_zv(all_predictors())

wine_wflow_reduce <- workflow() %>%
  add_model(wine_spec) %>%
  add_recipe(best_AIC_rec)

wine_fit_reduce <- wine_wflow_reduce %>%
  fit(data = wine_train)

kable(tidy(wine_fit_reduce), digits = 3)
```

After deciding that we are going to stick with the best AIC model, we
decide to remove citric acid, due to its high p-value, to increase our
model's accuracy.

$\log\Big(\frac{\hat\pi}{1-\hat\pi}\Big) = 407.674 + 0.472 \times fixed~acidity - 3.397 \times volatile~acidity + 0.214 \times residual~sugar - 7.505 \times chlorides + 0.007 \times free~sulfur~dioxide - 427.514 \times density + 2.289 \times pH + 2.515 \times sulphates + 0.462 \times alcohol - 1.228 \times colorwhite$

#### Cross Validation

We finally conducted cross validation, a technique for assessing how the
statistical analysis generalizes to the data set. It evaluates the
regression model by training several models on subsets (folds) of the
data set (we previously split into training and testing sets) and
evaluating them on the complementary subset of the data. To perform the
cross validation, we split the training data into 10 folds. After
performing the cross validation, we graphed metrics of accuracy and
ROC-AUC values of each individual fold for the two models on a
corresponding graph.

```{r, echo=FALSE}
set.seed(345)
folds <- vfold_cv(wine_train, v = 10)

wine_full_rs <- wine_wflow_full2 %>%
  fit_resamples(folds)

wine_AIC_rs <- wine_wflow_reduce %>%
  fit_resamples(folds)
```

```{r, echo=FALSE}
metrics_full <- collect_metrics(wine_full_rs, summarize = FALSE) %>% mutate(model = "Model 1 - Full")
metrics_reduce <- collect_metrics(wine_AIC_rs, summarize = FALSE) %>% mutate(model = "Model 2 - Reduce")
```

```{r, echo=FALSE, fig.height=4, fig.width=8, fig.align='center'}
metrics <- bind_rows(metrics_full, metrics_reduce) %>%
arrange(.metric)
acc_p <- ggplot(metrics %>% filter(.metric == "accuracy"), 
                aes(x = id, y = .estimate,
                group = model, color = model)) + geom_point() + geom_line() + labs(title = "Accuracy") + theme_bw()

aoc_p <- ggplot(metrics %>% filter(.metric == "roc_auc"), aes(x = id, y = .estimate,
    group = model, color = model)) + geom_point() + geom_line() + labs(title = "ROC-AUC") + theme_bw()

plot_grid(acc_p, aoc_p, nrow = 2)
```

The two models have similar accuracy and the ROC-AUC values for the 10
folds, so due to the principles of parsimony (i.e. model that provides
simpler explanation and has fewer terms), we prefer the more reduced one
or the best AIC model fit. Thus, the best AIC model is our final model.

#### Limitations

```{r, include=FALSE}
emplogitplot1(good_wine ~ c_fixed.acidity, 
              data = wine, 
              ngroups = 20)

emplogitplot1(good_wine ~ c_volatile.acidity, 
              data = wine, 
              ngroups = 20)

emplogitplot1(good_wine ~ c_free.sulfur.dioxide, 
              data = wine, 
              ngroups = 30)

emplogitplot1(good_wine ~ c_residual.sugar, 
              data = wine, 
              ngroups = 30)

emplogitplot1(good_wine ~ c_chlorides, 
              data = wine, 
              ngroups = 30)

emplogitplot1(good_wine ~ c_total.sulfur.dioxide, 
              data = wine, 
              ngroups = 30)

emplogitplot1(good_wine ~ c_density, 
              data = wine, 
              ngroups = 30)

emplogitplot1(good_wine ~ c_pH + color, 
              data = wine, 
              ngroups = 30)

emplogitplot1(good_wine ~ c_sulphates, 
              data = wine, 
              ngroups = 30)

emplogitplot1(good_wine ~ c_alcohol + color, 
              data = wine, 
              ngroups = 20)
```

When conducting a logistic regression, it is important to check the
conditions are satisfied. When evaluating if the log-odds have a linear
relationship with the predictors, we find that the conditions for
linearity are not satisfied for all the variables. The linearity
conditions are particularly not fulfilled for the variables free sulfur
dioxide (the spread takes on a curved shape), total sulfur dioxide (the
spread has similarly curved shape), and sulphates (the data points
appear to be randomly dispersed).

The independence conditions, however, are satisfied. The Vinho Verde
region is a vast region spreading 15500 hectares of vineyards in
far-north Portugal allowing the observations to be collected
independently from each other and the accumulated data are not collected
across an extended period of time.

### Results

Our final model is the reduced Best AIC model. To achieve this
conclusion, we first conducted a pairwise test to obtain a Best AIC
Model with no interactive terms, which is our reduced model. Then we
include the interactive terms and build another logistic model, as our
full model. We compared these two models' ROC curves, AIC, BIC. We also
check the two models' accuracy and ROC-AUC through cross-validation. We
found out that the two models have similar performances when predicting
the quality of wine. Due to the principles of parsimonious, we prefer
the simpler model. Therefore, we choose the reduced Best AIC Model as
our best model.

$\log\Big(\frac{\hat\pi}{1-\hat\pi}\Big) = 407.674 + 0.472 \times fixed~acidity - 3.397 \times volatile~acidity + 0.214 \times residual~sugar - 7.505 \times chlorides + 0.007 \times free~sulfur~dioxide - 427.514 \times density + 2.289 \times pH + 2.515 \times sulphates + 0.462 \times alcohol - 1.228 \times colorwhite$

From this model, we can see that some influential factors might be
chlorides, pH, sulphates, and volatile acidity. For example, the
coefficient for pH is 2.289, which means that for every 1-point increase
in pH, we expect the odds of wine being good to increase by a factor of
exp(2.289) = 9.87, holding all else constant.

To assess the model's predictive abilities, we fit it to test data. The
following ROC curve plot shows that the Best AIC Model does fairly well
in terms of predicting good wine vs bad wine.

```{r, echo=FALSE, fig.width=3, fig.height=2, message=FALSE, warning=FALSE, style= "float:right;position: relative; top: -80px;"}
best_AIC_pred <- best_AIC_fit %>% mutate(.pred_1 = exp(...1) / (1 + exp(...1))) %>% mutate(.pred_0 = 1 - .pred_1)
best_AIC_pred %>% roc_curve(truth = as.factor(good_wine), .pred_0) %>%
  ggplot(aes(x = 1 - specificity, y = sensitivity)) +
  geom_line() + labs(title = "Best AIC Model's ROC-AUC Curve")
```

### Discussion & Conclusion

From our analysis, we found that the variables that contribute
significantly to wine quality were fixed acidity, volatile acidity,
residual sugar, chlorides, free sulfur dioxide, total sulfur dioxide,
density, pH, and sulfates. Variables with biggest coefficients are
chlorides, volatile acidity, pH and sulphates. It is worth noting that
although the coefficient of density is very high, wine's density varies
between approximate 0.99 to 1, therefore it is impossible for wine'
density to jump 1 unit. So we did not include density as a influential
factor.

Some of the variables that negatively affected wine quality were
chlorides and volatile acidity (meaning when these variables increased,
quality was less likely to be good). Scientific reasons for this could
be that chlorides are a major contributor to saltiness, and volatile
acidity refers to the amount of acetic acid in wine, which is a kind of
acid which makes wine tastes like vinegar. Some variables that
positively affected wine quality were pH and sulphates. This could be
because wine with low pH tastes tart and crisp, while higher pH adds to
wine's ripeness. Additionally, sulphates are a food preservative widely
used in winemaking which protects wine against oxidation and maintains
the flavor and freshness of wine.

The result shows that, under wine experts' standards, wine that tastes
more salty and vinegary earns lower grades; on the other hand, wine
taste riper, mellower, fresher receive higher grades. This results is
coherent to our prediction. In the future, to improve the accuracy of
the model, different stepwise algorithms (p-values for example) can be
evaluated and adjusted. There are many different methods such as through
feature engineering, viewing potential interaction terms, or other
performance measurements such as machine learning algorithms to better
predict our results. Furthermore, a limitation of our dataset is the
unbalanced data, in which bad wine greatly outnumbers good wine. Thus,
another potential step in future work could be assessing models that
oversample good wine or downsample bad wine in order to address class
unbalance, and compare these models to our Best AIC model.

Analytical methods in the past century have permitted researchers to
gain detailed knowledge about the sugars and acids present in the
wine-making operation. Analysis and control of oxidation through
measuring both oxygen concentrations and sulfur dioxide concentrations
in grape juice and wine have resulted in significant improvements in
wine quality. Our results encourage further development in research to
measuring sulfur dioxide and the pH of grapes.

Quality in any wine is a function of the potential quality in the grape
and the conversion of grapes to wine. From our analysis, acidity is
particularly important for clean, fruity, balanced tastes. Low acidity
is also undesirable because it often leads to problems with fermentation
and preservation of the wine which manifest themselves as undesirable
tastes in the final wine.

Moreover analytical methods for sulfur dioxide determination can be
improved by research for wineyards such as techniques for measuring
phenols and enzymes or more precise methods of measuring acids of grapes
and wines. Sulfur dioxide controls oxidation in wines, but too much
gives the wine a very unpleasant flavor. Techniques for measuring
phenols and enzymes are still being developed in enology laboratories,
as are more precise methods of measuring residual sugars and the acids
of grapes and wines. 

Many analytical methods for sulfur dioxide determination and other
factors in wine quality are quite outdated but our research can be used
as an impetus to fund and significantly improve research in the specific
characteristics of wine quality.

### References

Paulo Cortez, University of Minho, Guimarães, Portugal,
http://www3.dsi.uminho.pt/pcortez A. Cerdeira, F. Almeida, T. Matos and
J. Reis, Viticulture Commission of the Vinho Verde Region(CVRVV), Porto,
Portugal, 2009.
